\section{Fundamental Region Reduction}

PyTex currently distinguishes between two reduction problems:

\begin{itemize}
  \item reducing crystal directions into class-specific inverse-pole-figure sectors
  \item reducing orientations into a symmetry-equivalent representative in orientation space
\end{itemize}

\subsection{Direction Reduction}

Let \(\mathbf{d}_c\) be a unit crystal direction and let \(\mathcal{C}\) denote the crystal symmetry group. PyTex forms the symmetry orbit
\[
  \mathcal{O}(\mathbf{d}_c) = \{ c \mathbf{d}_c \; | \; c \in \mathcal{C} \}.
\]

Direction reduction is then implemented by deterministic selection of the representative lying inside the supported direction-space fundamental sector
\[
  \mathbf{d}_{\mathrm{FR}} \in \mathcal{F}_{\mathrm{dir}}.
\]

This is the reduction used for inverse pole figures and IPF color keys.

\subsection{Orientation Reduction}

Let \(g\) denote an orientation mapping crystal coordinates into specimen coordinates. With crystal symmetry \(\mathcal{C}\) and optional specimen symmetry \(\mathcal{S}\), PyTex forms the orbit
\[
  \mathcal{O}(g) = \{ s g c \; | \; s \in \mathcal{S},\; c \in \mathcal{C} \}.
\]

PyTex currently supports two selection modes.

\paragraph{Exact orbit reduction in the quaternion hemisphere.}
If no reference orientation is supplied, each candidate is canonicalized into the unit-quaternion hemisphere and PyTex selects the symmetry-equivalent representative with maximal scalar part:
\[
  q_0 = \max_{q' \in \mathcal{O}(q)} q'_0.
\]
Because the rotation angle is
\[
  \omega = 2 \arccos(q_0)
\]
on that hemisphere, this is exactly the minimum-angle representative relative to the identity. PyTex then applies a deterministic lexicographic tie-break for orbit points lying on symmetry bisectors.

\paragraph{Reference-aware projection.}
If a reference orientation \(g_{\mathrm{ref}}\) is supplied, PyTex selects the candidate minimizing the unsymmetrized orientation distance to that reference.

\subsection{Current Limits}

\begin{itemize}
  \item PyTex now implements the exact orbit-reduction rule for the supported proper point groups already generated in the codebase, and the implementation is regression-tested class by class across that supported set.
  \item What remains ahead is a class-by-class closed-form boundary catalog and broader external parity fixtures for those exact boundaries.
  \item Current parity hardening therefore treats orientation projection as scientifically strong, but not yet exhaustively benchmarked against every external boundary reference.
\end{itemize}
