\section{Multiphase EBSD Graph Workflows}

This note records the current PyTex algorithmic contract for graph-backed and multiphase EBSD map
workflows.

\subsection{Scope}

The current implementation supports:

\begin{itemize}
  \item regular-grid adjacency through deterministic four- and eight-neighbor construction,
  \item irregular coordinate graphs through a distance-based neighborhood radius inferred from step
    size or the minimum positive coordinate spacing,
  \item phase-resolved crystal maps with explicit per-point phase identifiers, and
  \item phase-aware KAM and segmentation that preserve topology but exclude cross-phase
    misorientation pairs from within-phase metrics.
\end{itemize}

\subsection{Neighbor Doctrine}

For a crystal map with point set $\{x_i\}$, PyTex constructs an undirected neighbor graph
$\mathcal{G} = (\mathcal{V}, \mathcal{E})$.

\begin{itemize}
  \item On regular grids, $\mathcal{E}$ is defined by the requested connectivity and neighbor
    order.
  \item On irregular coordinates, $\mathcal{E}$ is defined by a radius threshold derived from the
    base spacing and the requested order.
\end{itemize}

This keeps neighborhood semantics explicit and allows KAM, grain segmentation, and boundary
extraction to share the same adjacency substrate.

\subsection{Phase Masking}

For phase-resolved maps, each point carries a phase identifier $p_i$. PyTex preserves the full
neighbor graph but restricts within-phase angular metrics to pairs satisfying

\[
p_i = p_j.
\]

This means the geometry remains visible while phase-incompatible misorientation calculations are not
silently mixed into KAM or grain-union logic.

\subsection{Texture Extraction}

Texture outputs from a multiphase crystal map require an explicit phase selector. PyTex does not
collapse a multiphase EBSD map into a single texture object without an explicit phase decision by
the user.

\subsection{Current Limits}

\begin{itemize}
  \item Irregular-coordinate graph construction is geometric rather than confidence-weighted.
  \item Majority-vote smoothing remains a regular-grid label operation.
  \item Multiphase EBSD import normalization is now supported, but detector-pattern semantics are
    still outside the current normalized dataset contract.
\end{itemize}
